---
title: "Final project - Data Processes UPM 2019/2020"
author: "Eduard Aymerich, Fernando Alonso, Rafael Timermans, Bowen Huang y Jaime Acosta"
date: "12/16/2019"
output: ioslides_presentation
---

# Movies analysis

```{r setup, include=FALSE}
# Load packages
library("ggplot2")
library("ggpubr")
library("corrplot")
library("Hmisc")
library("dplyr")
library("vcd")
library("tidyr")
library("tidyverse")  #data manipilation
library("GGally")     # nice scatterplot matrix
library("FactoMineR") # PCA computation
library("factoextra") # nice plotting for PCA objects
library("gridExtra")  # to build grid of plots
library("plyr")       # join function

# Load data set
tmdb_movies <- read.csv("./data/tmdb_5000_movies.csv", header = TRUE, sep = ",", fill = TRUE)
```

## Abstract
--

## Introduction and Related work
Cinema is the most complex and powerful art form in the present world. It can help us to better
understand our own lives, the lives of those around us and even how our society and culture operate. They can
shed light on political and spiritual matters too and can provide catharsis and perspective and may open our
eyes to new ways of thinking, feeling and pursuing our lives. This is why, in our way, we want to pay tribute with this movies analysis. Why we love some movies and hate others? We will conduct a study to try to understand better how movies and people are really connected.

We have found some inspirational data science projects related to this domain:

- [Data science analysis of movies released in the cinema between 2000 and 2017](https://medium.com/datadriveninvestor/data-science-analysis-of-movies-released-in-the-cinema-between-2000-and-2017-b2d9e515d032 "Data science analysis of movies released in the cinema between 2000 and 2017"): Consists in a general analysis of the data available on the IMDb website for movies released in the cinema between 2000 and 2017.

- [The most influential factor of imdb movie rating part in data scraping](https://towardsdatascience.com/the-most-influential-factor-of-imdb-movie-rating-part-i-data-scraping-61dc0c4dd518 "The most influential factor of imdb movie rating part in data scraping"): Use data-scraping technique to extract data from a IMDb movie list, create a Dataset and figures out what are the primary factors that influence a movie’s success.

-  [What makes a successful film predicting a films revenue and user rating with machine learning?](https://towardsdatascience.com/what-makes-a-successful-film-predicting-a-films-revenue-and-user-rating-with-machine-learning-e2d1b42365e7 "What makes a successful film predicting a films revenue and user rating with machine learning?"): It stroves to find out whether, knowing only things you could know before a film was released, what the rating and revenue of the film would be. It uses machine learning in a model on a randomized 90% of the movies, and then tested it on the remaining 10%.

- [Analyzing Movie Scores on IMDb and Rotten Tomatoes](http://rstudio-pubs-static.s3.amazonaws.com/336722_2193716117584b63a2a6ebb837217d85.html "Analyzing Movie Scores on IMDb and Rotten Tomatoes"): This project aims to obtain insights regarding the interpretation and comparison among these different scoring systems through data analysis and visualization. One focus in this project would be to visualize the distributions of scores and see how they vary across the years and across different genres. 

- [Movie Review Analysis: Emotion Analysis of IMDb Movie Reviews](https://www.researchgate.net/publication/305222147_Movie_Review_Analysis_Emotion_Analysis_of_IMDb_Movie_Reviews "Movie Review Analysis: Emotion Analysis of IMDb Movie Reviews"): Currently, moviegoers base their decisions as to which movie to watch by looking at the ratings of movies as well as reading some of the reviews at IMDb or Amazon. This paper argues that there is a better way: reviewers movie scores and reviews can be analyzed with respect to their emotion content, aggregated and projected onto a movie, resulting in an emotion map for a movie.


## Exploratory Data Analysis
In our analysis we will use the TMDB Movie Metadata dataset collected by Kaggle. The data was extracted from [The Movie Database](https://www.themoviedb.org) API to get almost 5000 films on 2017.

### TMDB Movie Metadata
This dataset contains 4803 rows with a total of 20 columns.

| *Name*               | *Data type*   | *Description*                                                   |
| :------------------: | :-----------: | :-------------------------------------------------------------: |
| Budget               | Integer       | Movie budget in dollars                                         |
| Genres               | String        | A comma-separated list of genres used to classify the film      |
| Homepage             | Url           | Official website of the film                                    |
| Id                   | Integer       | Identification number of the film created by TMBD               |
| Keywords             | String        | A comma-separated list of keywords used to classify the film    |
| Original_language    | String        | Original language of the film                                   |
| Original_title       | String        | Original title of the film                                      | 
| Overview             | String        | Short description of the film                                   |
| Popularity           | Decimal       | Popularity in the TMBD website                                  |
| Production_companies | String        | A comma-separated list of production companies in the film      |
| Production_countries | String        | A comma-separated list of production countries in the film      |
| Release_date         | Date          | Release date in YYYY-MM-DD format                               |
| Revenue              | Integer       | Movie revenue in dollars                                        |
| Runtime              | Integer       | The duration of the film in minutes                             |
| Spoken_languages     | String        | A comma-separated list of languages spoken in the original film |
| Status               | String        | Indicates if movie released. Values are "Released" or "Rumored" |
| Tagline              | String        | Short text to clarify or make you excited about the film        |
| Title                | String        | The title of the film                                           |
| Vote_average         | Decimal       | Average of users rating for the movie 0-10                      |
| Vote_count           | Integer       | Number of votes                                                 |

```{r, echo = TRUE}
summary(tmdb_movies)
```

The most relevant features for this study are Budget, Genres, Revenue, and vote_average. We will use more than these ones but are lees important.

We will clean a bit the data, so we will only use movies with all the fileds completed (no NA data), we wont consider either the movies with budget or revenue that equals 0, as we suppose that this is an error.
In addition, each film has multiple genres, but we will only consider the first genre of each movie.
```{r clean, include = FALSE}
# removing where we have no data to work with
movies2 <- na.omit(tmdb_movies)
view(movies2)
movies <-movies2

# theres a lot of films with revenue=0 and busget=0, we don'w't want this
movies2 <- movies2 %>% # 
  filter(budget>0) %>% #
  filter(revenue>0)

# genres are all in same column, baad, we take only the first genre (the most important i hope)
Genre <- movies2$genres
Genre <- as.data.frame(Genre)
Genre <- separate(Genre, col = Genre, into=c("1","2","3","4","5","6"))

movies2$genres <- Genre$`5`
movies2 <- na.omit(movies2)
# we have removed an extra film with no genre

# factor as factor and the column release date as date (spanish format)
movies2$genres = as.factor(movies2$genres)
movies2$production_countries = as.factor(movies2$production_countries)
movies2$status = as.factor(movies2$status)
```

#### Revenue and budget
First we will print a graph relating the revenue and budget, and with color information about the vote average, to see if we can se a relationship.

```{r revVSbudgVSvote, echo=FALSE}
# graph relations between budget and revenue with color as vote average
plot2 = ggplot(data=movies2, aes(x=budget, y=revenue))
plot2 + geom_jitter(aes(x=budget,y=revenue, color=vote_average), position=position_jitter(w=.3, h=.0))
# we dont see anything clear, 2 more graph with vote average
```


We can see that it may exist a relationship between the revenue of a film and the budget invested, but it doesn´t seem like any of those feature is related to vote average in the graph, so let´s print in a separate graph.

```{r relations with revenue, echo=FALSE}
# graph relations between vote average and revenue
plot3 = ggplot(data=movies2, aes(x=vote_average, y=revenue))
plot3 + geom_jitter(aes(x=vote_average, color=revenue), position=position_jitter(w=.3, h=.0)) -> p1

# graph relations between budget and vote average
plot4 = ggplot(data=movies2, aes(x=vote_average, y=budget))
plot4 + geom_jitter(aes(x=vote_average, color=budget), position=position_jitter(w=.3, h=.0)) -> p2

grid.arrange(p1, p2, ncol = 2)
```

We can see it clearer now, and we can appreciate that it is possible that there is a relationship between the vote average and the revenue, but doesn´t seem like that with the budget.

We are going to compare these features with the genre of the film, so we can see if action movies earn more mony than others genres movies.

#### Genre
Lets see if there is a relation between genre and revenue, so we can compare if a genre is better to make money than other.
```{r violin}
gen_vot <- ggplot(movies2, aes(x=genres, y=revenue),las=2) + 
  geom_violin()
gen_vot + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
We can see than Action, Drama and Science-Fiction make a lot more money than the others movie genres. To making a better comparision, we are now to clean a bit more the data. We are going to keep only the genres more popular (by revenue), and these are:
 - 1) Action
 - 2) Adventure
 - 3) Comedy
 - 4) Drama 
 - 5) Family
 - 6) Fantasy
 - 7) Romance
 - 8) Science-fiction
 - 9) Thriller
 
We are going to compare with the vote average too, so we will quit all the movies with few votes.
```{r violin2}
summary(movies2$vote_count)
```
There is a big difference between the most voted and the least, so we are not going to use the ones in the first quartile.

```{r clean_votes_and_genres, include= TRUE}
#there are a lot of films who doesnt have many votes (half of movies have less than 471 votes), we have to use only fimls which have more than 178 votes
#(1st quartile)
movies_reduced_votes <- movies2 %>% # data frame to start with
  filter(vote_count>178)  %>%
  filter(genres=="Action" | genres=="Adventure" | genres=="Animation" | genres=="Comedy" | genres=="Drama" | genres=="Family" | genres=="Fantasy" | genres=="Romance" | genres=="Thriller" | genres=="Science")
```
Once we have a cleaner dataset, we want to compare the genre with revenue and score. In order to do that, we convert the vote average in facotr, so we round the number (it is a 1 to 10, scale).

```{r facet_the_vote, include= TRUE}
# making a facet with vote integers
movies_reduced_votes$vote_average = sapply(movies_reduced_votes$vote_average, function(x) floor(x/1))
movies_reduced_votes$vote_average <- as.factor(movies_reduced_votes$vote_average)
```

We make a plot now comparing these three variables, to see if there is a pattern.
```{r plot_violin_by_vote}
# we only lost like 400 movies, they were not common films then, good
gen_vot <- ggplot(movies_reduced_votes, aes(x=genres, y=revenue)) + 
  geom_violin()

gen_vot <- gen_vot+facet_wrap(~vote_average)
gen_vot + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
We see than the genres revenue are similar, but when the score rounds the seven, action films earn much more money than the rest. To see if there is a different pattern with the action movies, we are going to do the same graphics than before but only with action movies.

#### Action movies
We need to clean the data first:
```{r clean_action}
# lets see revenue vs budget in action films compared to vote score, only for action movies
action_movies <- movies_reduced_votes %>%
  filter(genres == "Action")
```

And lastly we print the same graphs than before
```{r action_graphs, echo= FALSE}
# graph relations between busgete and revenue
plotAction = ggplot(data=movies2, aes(x=budget, y=revenue))
plotAction + geom_jitter(aes(x=budget, color=revenue), position=position_jitter(w=.3, h=.0))

gen_vot <- ggplot(action_movies, aes(x=budget, y=revenue)) + 
  geom_violin()
gen_vot <- gen_vot+facet_wrap(~vote_average)
gen_vot + theme(axis.text.x = element_text(angle = 90, hjust = 1))

# grid.arrange(p1, p2, ncol = 2)
```
Movies than have an 8 as a vote_average don't earn a lot or few money. We can also see that the ones that earn a lot of money they have an average budget.

The first graph is the same that the one considering all genres, so it is no usefull for the study. However, we can see that in action films it is possible that the revenue is indeed related with the vote_average, so as the vote average rise, there are more films which have more revenue. The exception is the movies with more than an 8. These movies doesn´t have a big revenue, but they doen´t invest too much money in the budget neither, so they are likely profitable.

#### Money vs Time
We are going to plot some graphs tto see if there is a tendency to invest or earn more money with time. We are going to consider all the movies.

```{r clean_date, echo=FALSE}
# compare revenue with year
movies2$release_date <- as.Date(movies2$release_date, "%Y-%m-%d")

p4 <- ggplot(movies2, aes(x=release_date, y=revenue)) +
  geom_point(color="steelblue") +
  xlab("")

p5 <- ggplot(movies2, aes(x=release_date, y=budget)) +
  geom_point(color="steelblue") +
  xlab("")

grid.arrange(p4, p5, ncol = 2)
```
Seems like indeed there is more money invested each year. To have a clearer view, we have to plot the graph with only the median value, not with all the films.

```{r mean_year, echo=FALSE}
# first mutate to obtain the mean, then plot.
movies3 <- movies2 %>%
  mutate( year = format(release_date, "%Y")) %>%
  group_by( year) %>%
  mutate(revenue_mean=mean(revenue)) %>%
  mutate(budget_mean=mean(budget)) 
# summarise(total = mean(revenue))

p7 <- ggplot(movies3, aes(x=release_date, y=revenue_mean)) +
  geom_point() +
  geom_line(color="steelblue")+
  xlab("")

p8 <- ggplot(movies3, aes(x=release_date, y=budget_mean)) +
  geom_point() +
  geom_line(color="steelblue")+
  xlab("")
  
grid.arrange(p7, p8, ncol = 2)
```
It is obvious that, despite of the exception in the 20s years, the budget increase each year. In the revenue is not that obvious and it is possible that we need to do a time series analysis to be sure, but the tendency is increasing too.


## Questions of interest
- What attributes makes the film better (with better we refer to the money earned, the popularity, the ratings)?
- Does the duration of film influence the average of users rating?
- Are some film genres more popular than others? Have it changed during years?
- Have old films better critic scores than recent ones?
- Can a film earn a more than 1 billion dollars if the genre is not one of the top3 popular gender in that year?
- The release year influences in the popularity and revenue of a movie?

## Methods

### Strength of relationships
First we will clean the data
```{r clean_data, echo = TRUE}
#### Remove budget higher from 0 and revenue higher than 0
tmdb_movies_clean <- tmdb_movies %>% 
                     filter(budget>0) %>% 
                     filter(revenue>0) 

#### Remove where there is no data
tmdb_movies_clean <- na.omit(tmdb_movies_clean)
```

Question 1) First we will see which attribute makes the film earn more money.

Step 1.1) Select rows that we will use. In this case the earn money (*revenue*) are related to the variables *budget*, *genres*, *production companies*, *release date* and *spoken languages*.
```{r revenue_related, echo = TRUE}
movies_related_revenue <- tmdb_movies_clean %>% select(revenue, budget, genres, popularity, production_companies, release_date, spoken_languages)
```

Step 1.2) Compare the variable *revenue* with other variables.
```{r compare_revenue_budget, echo = TRUE}
# Compare *revenue* with *budget*
movies_r <- movies_related_revenue %>% select(revenue, budget)
cor(movies_r$revenue, movies_r$budget)

ggplot(movies_related_revenue, aes(x=revenue, y=budget))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Revenue", y="Budget", title="Relation between Revenue and Badget")

regression_model = lm(formula = revenue ~ budget + genres + popularity + production_companies + release_date + spoken_languages, data=movies_related_revenue)

summary(regression_model)
```

Question 2) Secondly we will see which attribute makes the film more popular in the TMBD website.
```{r more_popular, echo = TRUE}
# Compare *popularity* with *revenue*
movies_r <- movies_related_revenue %>% select(revenue, popularity)
cor(movies_r$revenue, movies_r$popularity)

ggplot(movies_related_revenue, aes(x=revenue, y=popularity))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Revenue", y="Popularity", title="Relation between Revenue and Popularity")

regression_model = lm(formula = revenue ~ popularity + budget + genres + production_companies + release_date + spoken_languages, data=movies_related_revenue)

summary(regression_model)

# Compare *popularity* with *budget*

# Compare *popularity* with *genres*

```

Question 3) Now we will see which attribute makes the film better voted by users.
```{r better_voted, echo = TRUE}
# Compare *vote_average* with *revenue*

# Compare *vote_average* with *budget*

# Compare *vote_average* with *genres*

```

Question 4) After that we will see if the duration of film influence the average of users rating.
```{r duration_influence_user_rating, echo = TRUE}
# Numerical value of the correlation
movies_r <- tmdb_movies_clean %>% select(vote_average, runtime)
cor(movies_r$vote_average, movies_r$runtime)

# Graphical representation
ggplot(tmdb_movies_clean, aes(x=vote_average, y=runtime)) +
geom_point(shape=1) +
geom_smooth(method=lm)
```

Question 5) The next stage will be analyse the more popular movies genres and how has it changed during years.
```{r popular_genres_during_years, echo = TRUE}
# Top3 genres during years
# Select the top genre for each year
popular_genre <- tmdb_movies_clean
popular_genre$year <- format(as.Date(popular_genre$release_date, format="%Y-%m-%d"),"%Y") # add year column

# get first genre
genre <- popular_genre$genres
genre <- as.data.frame(genre)
genre <- separate(genre, col = genre, into = c("1","2","3","4","5","6"))
popular_genre$genres <- genre$`5`
popular_genre <- na.omit(popular_genre) # we have removed an extra film with no genre

popular_genre <- popular_genre %>%
                 select(genres, year)

# Subdivisions years
old_genre <- popular_genre %>% 
              filter(year<=1980)
count(old_genre, "genres")
# Top3:
# 1) Drama
# 2) Action and Comedy (equal number of films)

neutral_genre <- popular_genre %>% 
              filter(year>1980 && year<2015)
count(neutral_genre, "genres")
# Top3:
# 1) Drama
# 2) Comedy
# 3) Action

new_genre <- popular_genre %>% 
              filter(year>=2015)
count(new_genre, "genres")
# Top3:
# 1) Action
# 2) Drama
# 3) Comedy

```

Question 6) Moreover we will see if old films have better user scores than recent ones.
```{r old_films_better_scores, echo = TRUE}
# Compare films vote_average of films released on 1980 or before with the ones released on 2015 or after.
tmdb_movies_clean$year <- format(as.Date(tmdb_movies_clean$release_date, format="%Y-%m-%d"),"%Y")

movies_r <- tmdb_movies_clean %>% 
            select(vote_average, vote_count, year)

movies_old <- movies_r %>% 
              filter(year<=1980)
nrow(movies_old) # number of rows
mean(movies_old$vote_average) # average vote
sum(movies_old$vote_count) # number of votes

movies_now <- movies_r %>% 
              filter(year>=2015)
nrow(movies_now) # number of rows
mean(movies_now$vote_average) # average vote
sum(movies_now$vote_count) # number of votes
```

Question 7) Lastly we will ask if a film can earn a more than 1 billion dollars if the genre is not one of the most popular gender in that year.
```{r film_earn_billion, echo = TRUE}
# Filter movies with revenue more than 1.000.000.000 dollars
movies_more_revenue <- tmdb_movies_clean %>% 
                       filter(revenue>1000000000)

# Select years from dates of the movies with more revenue
movies_more_revenue$year <- format(as.Date(movies_more_revenue$release_date, format="%Y-%m-%d"),"%Y")

# get first genre
genre <- movies_more_revenue$genres
genre <- as.data.frame(genre)
genre <- separate(genre, col = genre, into = c("1","2","3","4","5","6"))
movies_more_revenue$genres <- genre$`5`
movies_more_revenue <- na.omit(movies_more_revenue) # we have removed an extra film with no genre

movies_more_revenue <- movies_more_revenue %>% select(revenue, year, genres)
movies_more_revenue <- movies_more_revenue[order(movies_more_revenue$year),]
names(movies_more_revenue)[3] <- "movie_genre"

# Select the top genre for each year
year_top_genre <- tmdb_movies_clean
year_top_genre$year <- format(as.Date(year_top_genre$release_date, format="%Y-%m-%d"),"%Y") # add year column

# get first genre
genre <- year_top_genre$genres
genre <- as.data.frame(genre)
genre <- separate(genre, col = genre, into = c("1","2","3","4","5","6"))
year_top_genre$genres <- genre$`5`
year_top_genre <- na.omit(year_top_genre) # we have removed an extra film with no genre

# get genre and year
year_top_genre <- year_top_genre %>% select(genres, year) 

year_genre <- year_top_genre %>% select(year, genres) %>% 
              group_by(year)
year_genre

# Compare the genre of each film with the top genre of the release year
join(movies_more_revenue, year_genre, by ="year")
```

Question 8) Finally we will see if the release year influences in the popularity and revenue of a movie.
```{r release_year_influences_popularity, echo = TRUE}
tmdb_movies_clean$year <- format(as.Date(tmdb_movies_clean$release_date, format="%Y-%m-%d"),"%Y")

movies_r <- tmdb_movies_clean %>% 
            select(popularity, revenue, year)
movies_r
```

### Prediction

For the pre-processing stage, we have made some transformations:
 - Remove irrelevant variables, such as name (the variable id identifies each entry), keywords and some variables, which format transformation exceeds the scope of this project.
- Convert the name and description attributes, that could be interpreted as categorical variables, to a new field which contains the length of its name/description, in order to find fi this length is relevant for the model.
- Delete missing values for target variables, such as revenue, vote or popularity, in order to train the model correctly.
The final data frame looks like this:

``
     budget original_language popularity release_date    revenue runtime   status vote_average vote_count name_length tagline_length
1 237000000                en  150.43758   2009-12-10 2787965087     162 Released          7.2      11800           6             27
2 300000000                en  139.08262   2007-05-19  961000000     169 Released          6.9       4500          40             46
3 245000000                en  107.37679   2015-10-26  880674609     148 Released          6.3       4466           7             21
4 250000000                en  112.31295   2012-07-16 1084939099     165 Released          7.6       9106          21             15
5 260000000                en   43.92699   2012-03-07  284139100     132 Released          6.1       2124          11             36
6 258000000                en  115.69981   2007-05-01  890871626     139 Released          5.9       3576          12             18
``

## Results

### Strength of relationships

### Prediction

## Discussion and Future work


