

## Exploratory Data Analysis
In our analysis we will use 2 different datasets. The first one, called IMDB Data, was collected by Kaggle on 2016. The data was extracted by querying the REST Api of [IMDB](https://www.imdb.com) to get the best 1000 films. The second one, dataset, called TMDB Movie Metadata, was also collected by Kaggle. The data was extracted from [The Movie Database](https://www.themoviedb.org) API to get 5000 films on 2017.


### TMDB Movie Metadata

```{r message= FALSE, warning= FALSE, include= FALSE}
setwd("D:/AplicacionesInformaticas/workspaceR")

# Load up ggplot2 package to use for visualizations and dplyr for data manipulation
library(ggplot2)
library(dplyr)
library("vcd")
library(tidyr)

library(tidyverse) #data manipilation
library(GGally) # nice scatterplot matrix
library(FactoMineR) # PCA computation
library(factoextra) # nice plotting for PCA objects
library(gridExtra) # to build grid of plots

# Load data set
# titanic <- read.csv("Asignaturas/IDA/movies.RData", header = TRUE)
movies <- read.csv(file="./datasets/tmdb_5000_movies.csv", header=TRUE, sep=",")
```

This dataset contains 4803 rows with a total of 20 columns.

| *Name*               | *Data type*   | *Description*                                                   |
| :------------------: | :-----------: | :-------------------------------------------------------------: |
| Budget               | Integer       | Movie budget in dollars                                         |
| Genres               | String        | A comma-separated list of genres used to classify the film      |
| Homepage             | Url           | Official website of the film                                    |
| Id                   | Integer       | Identification number of the film created by TMBD               |
| Keywords             | String        | A comma-separated list of keywords used to classify the film    |
| Original_language    | String        | Original language of the film                                   |
| Original_title       | String        | Original title of the film                                      | 
| Overview             | String        | Short description of the film                                   |
| Popularity           | Decimal       | Popularity in the TMBD website                                  |
| Production_companies | String        | A comma-separated list of production companies in the film      |
| Production_countries | String        | A comma-separated list of production countries in the film      |
| Release_date         | Date          | Release date in YYYY-MM-DD format                               |
| Revenue              | Integer       | Movie revenue in dollars                                        |
| Runtime              | Integer       | The duration of the film in minutes                             |
| Spoken_languages     | String        | A comma-separated list of languages spoken in the original film |
| Status               | String        | Indicates if movie released. Values are "Released" or "Rumored" |
| Tagline              | String        | Short text to clarify or make you excited about the film        |
| Title                | String        | The title of the film                                           |
| Vote_average         | Decimal       | Average of users rating for the movie 0-10                      |
| Vote_count           | Integer       | Number of votes                                                 |


```{r movies, echo = FALSE, include= FALSE}
str(movies)
```

The most relevant features for this study are Budget, Genres, Revenue, and vote_average. We will use more than these ones but are lees important.



We will clean a bit the data, so we will only use movies with all the fileds completed (no NA data), we wont consider either the movies with budget or revenue that equals 0, as we suppose that this is an error.
In addition, each film has multiple genres, but we will only consider the first genre of each movie.
```{r clean, include = FALSE}
# view(movies)
#quitamos donde no hay datos para trabajar mejor
movies2 <- na.omit(movies)
view(movies2)
movies<-movies2
#theres a lot of films with revenue=0 and busget=0, we donw want this
# 
movies2 <- movies2 %>% # 
  filter(budget>0) %>% #
  filter(revenue>0)

#genres are all in same column, baad, we take only the first genre (the most important i hope)
Genre<-movies2$genres
Genre<-as.data.frame(Genre)
Genre<-separate(Genre, col = Genre, into=c("1","2","3","4","5","6"))
##View(Genre)
movies2$genres<-Genre$`5`
movies2 <- na.omit(movies2)
#we have removed an extra film with no genre

# factor as factor and the column release date as date (spanish format)
movies2$genres=as.factor(movies2$genres)
#movies2$original_languaje=as.factor(movies2$original_languaje)
movies2$production_countries=as.factor(movies2$production_countries)
#movies2$spoken_languajes=as.factor(movies2$spoken_languajes)
movies2$status=as.factor(movies2$status)
```

#### Revenue and budget
First we will print a graph relating the revenue and budget, and with color information about the vote average, to see if we can se a relationship.

```{r revVSbudgVSvote, echo=FALSE}
#graph relations between budget and revenue with color as vote average
plot2=ggplot(data=movies2, aes(x=budget, y=revenue))
plot2+geom_jitter(aes(x=budget,y=revenue, color=vote_average), position=position_jitter(w=.3, h=.0))
#we dont see anything clear, 2 more graph with vote average
```


We can see that it may exist a relationship between the revenue of a film and the budget invested, but it doesn´t seem like any of those feature is related to vote average in the graph, so let´s print in a separate graph.

```{r relations with revenue, echo=FALSE}
#graph relations between vote average and revenue
plot3=ggplot(data=movies2, aes(x=vote_average, y=revenue))
plot3+geom_jitter(aes(x=vote_average, color=revenue), position=position_jitter(w=.3, h=.0)) -> p1

#graph relations between budget and vote average
plot4=ggplot(data=movies2, aes(x=vote_average, y=budget))
plot4+geom_jitter(aes(x=vote_average, color=budget), position=position_jitter(w=.3, h=.0)) -> p2

grid.arrange(p1, p2, ncol = 2)
```

We can see it clearer now, and we can appreciate that it is possible that there is a relationship between the vote average and the revenue, but doesn´t seem like that with the budget.

We are going to compare these features with the genre of the film, so we can see if action movies earn more mony than others genres movies.

####Genre
Lets see if there is a relation between genre and revenue, so we can compare if a genre is better to make money than other.
```{r violin}
gen_vot <- ggplot(movies2, aes(x=genres, y=revenue),las=2) + 
  geom_violin()
gen_vot + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
 We can see than Action, Drama and Science-Fiction make a lot more money than the others movie genres. To making a better comparision, we are now to clean a bit more the data. We are going to keep only the genres more popular (by revenue), and these are:
 - Action
 - Adventure
 - Comedy
 - Drama 
 - Family
 - Fantasy
 - omance
 - Science-fiction
 - Thriller
 
 We are going to compare with the vote average too, so we will quit all the movies with few votes.
 
```{r violin2}
summary(movies2$vote_count)
```
There is a big difference between the most voted and the least, so we are not going to use the ones in the first quartile.

```{r clean_votes_and_genres, include= TRUE}
#there are a lot of films who doesnt have many votes (half of movies have less than 471 votes), we have to use only fimls which have more than 178 votes
#(1st quartile)
movies_reduced_votes <- movies2 %>% # data frame to start with
  filter(vote_count>178)  %>%
  filter(genres=="Action" | genres=="Adventure" | genres=="Animation" | genres=="Comedy" | genres=="Drama" | genres=="Family" | genres=="Fantasy" | genres=="Romance" | genres=="Thriller" | genres=="Science")
```
Once we have a cleaner dataset, we want to compare the genre with revenue and score. In order to do that, we convert the vote average in facotr, so we round the number (it is a 1 to 10, scale).

```{r facet_the_vote, include= TRUE}
#making a facet with vote integers
movies_reduced_votes$vote_average = sapply(movies_reduced_votes$vote_average, function(x) floor(x/1))
movies_reduced_votes$vote_average<-as.factor(movies_reduced_votes$vote_average)
```

we make a plot now comparing these three variables, to see if there is a pattern.
```{r plot_violin_by_vote}
#we only lost like 400 movies, they were not common films then, good
gen_vot <- ggplot(movies_reduced_votes, aes(x=genres, y=revenue)) + 
  geom_violin()

gen_vot<-gen_vot+facet_wrap(~vote_average)
gen_vot+ theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

We see than the genres revenue are similar, but when the score rounds the seven, action films earn much more money than the rest. To see if there is a different pattern with the action movies, we are going to do the same graphics than before but only with action movies.

#### Action movies
We need to clean the data first:

```{r clean_action}
#lets see revenue vs budget in action films compared to vote score, only for action movies
action_movies <- movies_reduced_votes %>%
  filter(genres=="Action")

```

And lastly we print the same graphs than before

```{r action_graphs, echo= FALSE}
#graph relations between busgete and revenue
plotAction=ggplot(data=movies2, aes(x=budget, y=revenue))
plotAction+geom_jitter(aes(x=budget, color=revenue), position=position_jitter(w=.3, h=.0))

gen_vot <- ggplot(action_movies, aes(x=budget, y=revenue)) + 
  geom_violin()
gen_vot<-gen_vot+facet_wrap(~vote_average)
gen_vot+ theme(axis.text.x = element_text(angle = 90, hjust = 1))

#grid.arrange(p1, p2, ncol = 2)
# wow, las que tienen un 8 ni gastan mucho ni ganan poco, y seguimos comprobando que ganan mucho las que tienen un budget medio
```

The first graph is the same that the one considering all genres, so it is no usefull for the study. However, we can see that in action films it is possible that the revenue is indeed related with the vote_average, so as the vote average rise, there are more films which have more revenue. The exception is the movies with more than an 8. These movies doesn´t have a big revenue, but they doen´t invest too much money in the budget neither, so they are likely profitable.

#### Money vs Time
We are going to plot some graphs tto see if there is a tendency to invest or earn more money with time. We are going to consider all the movies.

```{r clean_date, echo=FALSE}
####################### compare revenue with year
movies2$release_date <- as.Date(movies2$release_date, "%Y-%m-%d")
#head(movies2)

p4 <- ggplot(movies2, aes(x=release_date, y=revenue)) +
  geom_point(color="steelblue") +
  xlab("")


p5 <- ggplot(movies2, aes(x=release_date, y=budget)) +
  geom_point(color="steelblue") +
  xlab("")

grid.arrange(p1, p2, ncol = 2)
```

Seems like indeed there is more money invested each year. To have a clearer view, we have to plot the graph with only the median value, not with all the films.

```{r mean_year, echo=FALSE}
#first mutate to obtain the mean, then plot.
movies3 <- movies2 %>%
  mutate( year = format(release_date, "%Y")) %>%
  group_by( year) %>%
  mutate(revenue_mean=mean(revenue)) %>%
  mutate(budget_mean=mean(budget)) 
 # summarise(total = mean(revenue))

p7 <- ggplot(movies3, aes(x=release_date, y=revenue_mean)) +
  geom_point() +
  geom_line(color="steelblue")+
  xlab("")


p8 <- ggplot(movies3, aes(x=release_date, y=budget_mean)) +
  geom_point() +
  geom_line(color="steelblue")+
  xlab("")
grid.arrange(p7, p8, ncol = 2)
```

It is obvious that, despite of the exception in the 20s years, the budget increase each year. In the revenue is not that obvious and it is possible that we need to do a time series analysis to be sure, but the tendency is increasing too.